The Semantics and Implementation of \CEU: a Synchronous Reactive Language based on Esterel

### 2015-09-24 paper.tex
Revisao Secao 2.1

\begin{figure}
\begin{minipage}[t]{0.48\linewidth}
\begin{lstlisting}[numbers=left,xleftmargin=3.5em,mathescape=true]
// ESTEREL
loop
   abort
      [
         await A
      $\|$
         await B
      ];
      emit O
   when R
end
\end{lstlisting}
\end{minipage}
%
\begin{minipage}[t]{0.48\linewidth}
\begin{lstlisting}[numbers=left,xleftmargin=3.5em]
// CEU
loop do
   par/or do
      par/and do
         await A;
      with
         await B;
      end
      emit O;
   with
      await R;
   end
end
\end{lstlisting}
\end{minipage}
%\rule{8.4cm}{0.37pt}
\caption{ A control specification implemented in Esterel and \CEU:
\emph{``Emit \code{O} after \code{A} and \code{B}, resetting each \code{R}.''}
%\newline
{\small
%TODO.
}
\label{lst.abro}
}
\end{figure}

@@

Another distinction regarding event handling is that Esterel supports 
same-cycle bi-directional communication~\cite{esterel.compiling}, i.e., two 
threads can react to one another during the same cycle due to mutual signal 
dependency.
%
\CEU imposes a restriction that an \code{await} is only valid for the next 
reaction, i.e., if an \code{await} and \code{emit} for the same event occur 
simultaneously in parallel trails, the \code{await} does not awake.
%
These \emph{delayed awaits} avoid corner cases of instantaneous termination and 
re-execution of statements in the same reaction inside loops (known as 
\emph{schizophrenic statements})~\cite{esterel.schizo}.
%
The example in the left of Figure~\ref{lst.delay} (\code{CODE-1}) relies on 
this restriction to avoid infinite execution:
Although both sides of the \code{par/or} have an \code{await} statement, the 
\code{loop} would restart instantaneously if the \code{await} could awake in 
the same reaction that reaches it.
%
In atypical scenarios requiring immediate awake, delayed awaits can be 
circumvented by manually copying or making some transformation in the code to 
execute on awake.
%
From our experience, in some cases we need to execute a block of code 
periodically, \emph{including the current reaction}, from requests of an 
internal event, as illustrated in the middle of Figure~\ref{lst.delay} 
(\code{CODE-2}).
In this case, moving the \code{await} to the end of the loop suffices to 
execute the periodic code immediately (and then in reaction to each \code{emit} 
request).
If the periodic \code{emit} depends on a condition, the code becomes more 
intricate, requiring to copy the condition test to the periodic code, as 
illustrated in the right of Figure~\ref{lst.delay} (\code{CODE-3}).

@@

%Therefore, there is a single order of execution for side effects (represented 
%by \code{mem} operations), thus, reaction chains are always deterministic.
%
% TODO
%\footnote{We could extend the semantics to describe the full execution of a 
%program by holding new incoming external events in a queue and processing them 
%in consecutive reaction chains that never overlap.}
%
%The semantics requires an explicit stack to properly nest the emission of 
%internal events.
%to prevent that $awaiting$ expressions awake in the same reaction they are 
%reached.

@@

% TODO:, allowing \CEU{} to be easily embedded in platforms:

>>>>

\begin{figure}
\begin{minipage}[t]{0.48\linewidth}
\begin{lstlisting}[numbers=left,xleftmargin=3.5em,mathescape=true]
// ESTEREL
loop
   abort
      [
         await A
      $\|$
         await B
      ];
      emit O
   when R
end
\end{lstlisting}
\end{minipage}
%
\begin{minipage}[t]{0.48\linewidth}
\begin{lstlisting}[numbers=left,xleftmargin=3.5em]
// CEU
loop do
   par/or do
      par/and do
         await A;
      with
         await B;
      end
      emit O;
   with
      await R;
   end
end
\end{lstlisting}
\end{minipage}
%\rule{8.4cm}{0.37pt}
\caption{ A control specification implemented in Esterel and \CEU:
\emph{``Emit \code{O} after \code{A} and \code{B}, resetting each \code{R}.''}
\label{lst.abro}
}
\end{figure}

@@

Another distinction regarding event handling in comparison to \CEU is that 
Esterel supports same-cycle bi-directional 
communication~\cite{esterel.compiling}, i.e., two threads can react to one 
another during the same cycle due to mutual signal dependency.
%
\CEU imposes a restriction that an \code{await} to an internal event is only 
valid for the next reaction, i.e., if an \code{await} and \code{emit} occur 
simultaneously in parallel trails, the \code{await} does not awake.
%
These \emph{delayed awaits} avoid corner cases of instantaneous termination and 
re-execution of statements in the same reaction inside loops (known as 
\emph{schizophrenic statements})~\cite{esterel.schizo}.
%
The example in the left of Figure~\ref{lst.delay} (\code{CODE-1}) relies on 
this restriction to avoid infinite execution:
Although both sides of the \code{par/or} have an \code{await} statement, the 
\code{loop} would restart instantaneously if the \code{await} could awake in 
the same reaction that reaches it.
%
In atypical scenarios requiring immediate awake, delayed awaits can be 
circumvented by manually copying or making some transformation in the code to 
execute on awake.
%
From our experience, in some cases we need to execute a block of code 
periodically, \emph{including the current reaction}, starting from requests of 
an internal event, as illustrated in the middle of Figure~\ref{lst.delay} 
(\code{CODE-2}).
In this case, moving the \code{await} to the end of the loop suffices to 
execute the periodic code immediately (and then in reaction to each \code{emit} 
request).
If the periodic \code{emit} depends on a condition, the code becomes more 
intricate, requiring to copy the condition test to the periodic code, as 
illustrated in the right of Figure~\ref{lst.delay} (\code{CODE-3}).

@@

@@
